# Tips and Tricks

- Spell check functionality in RStudio

Although often overlooked, RStudio can check your spelling


- New line for each sentence

## Global chunk options

It can be useful to set chunk options globally for all (following) chunks to avoid retyping or copy-and-pasting

```{r eval = FALSE}
# Save all plots as 600 DPI TIFF-files
knitr::opts_chunk$set(dev = "tiff", dpi = 600)

# Do not evaluate subsequent chunks (debugging or fine-tuning)
knitr::opts_chunk$set(eval = FALSE)
```

See the [knitr chunk options and package options](https://yihui.name/knitr/options/) for an overview of settings

## Meaningful chunk names

...


## Text references

Because `papaja` extends `bookdown` you can use text references in any `papaja` document.
A text reference consists of a unique label---defined as `(ref:unique-label)` somewhere in the body of the document (not inside a code chunk)---and the text that the label stands in for.
Text references must be defined in a separate single-line paragraph with empty lines above and below:

```
(ref:my-caption) This is a caption for my table.
```

```{block2, text-references-warning, type='rmdcaution'}
The definition of a text reference must be on a single line and should not end with a white space.
Hence, the following will *not* work as expected

~~~
(ref:my-caption) 
This is a caption for my table.
~~~

```

Using text references for table and figure captions has several advantages:

1. Markdown and $\LaTeX$ syntax is not well supported in chunk options, such as `fig.cap`, or elsewhere inside a code chunk, e.g., in the `caption` argument of `apa_table()`.
Specifically, Markdown formatting and citation syntax are ignored (rendered as-is) and, for example, `\` and `_` must be escaped to prevent errors in either R or $\LaTeX$.
None of these limitations apply to text references.
2. Long captions can impair the readability of the document when they are part of chunk options or R code.
3. `knitr` considers modifications of the text in `fig.cap` as changes to the code chunk and hence may invalidate the cache of a chunk.
As a consequence cached results of the code chunk need to be recomputed.
Changes to the text of a text reference do not invalidate the cached computations.
4. It's straight forward to include inline code chunks in text references.

Text references can be used to duplicate information throughout the document.
Consider the following example.

~~~
(ref:aesthetics) Points represent conditions means, error bars represent 955% confidence intervals.

(ref:caption1) An interesting plot.

```{r fig.cap = paste("(ref:caption1)", "(ref:aesthetics)")}`r ''`
apa_beeplot(data = npk, id = "block", dv = "yield", factors = "N")
```

(ref:caption2) Another interesting plot.

```{r fig.cap = paste("(ref:caption2)", "(ref:aesthetics)")}`r ''`
apa_beeplot(data = npk, id = "block", dv = "yield", factors = "N")
```
~~~

The resulting figure caption combines the two text references, Figure\ \@ref(fig:text-reference-figure).

(ref:aesthetics2) Points represent conditions means, error bars represent 955% confidence intervals.

(ref:caption3) An interesting plot.

```{r text-reference-figure, fig.cap = paste("(ref:caption3)", "(ref:aesthetics2)"), echo = FALSE}
apa_beeplot(data = npk, id = "block", dv = "yield", factors = "N")
```

Because the information about what points and error bars represent is repeated using a text reference, and not by literal repetition throughout the document, it's easy to correct the typo (`955% confidence intervals`) and be sure that it is corrected in every instance.


## Useful RStudio Addins

- [`citr`](https://github.com/crsh/citr): Insert Markdown Citations
- [`remedy`](https://github.com/ThinkR-open/remedy): Keyboard shortcuts for Markdown formatting
- [`splitChunk`](https://github.com/LudvigOlsen/splitChunk): Split R Markdown code chunks
- [`gramr`](https://github.com/ropenscilabs/gramr): Write-good linter
- [`wordcount`](https://github.com/benmarwick/wordcountaddin): Word counts and readability statistics

Set up keyboard shortcuts via<br />`Tools > Modify keyboard shortcuts`


Suggested keyboard shortcuts

| Package | Addin | Keyboard shortcut |
|:--------|:------|:------------------|
| `citr` | Insert citation | <kbd>Shift</kbd> + <kbd>Alt</kbd>+<kbd>R</kbd> |
| `wordcount` | Word count | <kbd>Shift</kbd> + <kbd>Alt</kbd>+<kbd>C</kbd> |
| `splitChunk` | Chunk split | <kbd>Shift</kbd> + <kbd>Alt</kbd>+<kbd>S</kbd> |
| `remedy` | Bold | <kbd>Shift</kbd> + <kbd>Alt</kbd>+<kbd>B</kbd> |
|          | Italic | <kbd>Shift</kbd> + <kbd>Alt</kbd>+<kbd>I</kbd> |
|          | Backtick | <kbd>Shift</kbd> + <kbd>Alt</kbd>+<kbd>P</kbd> |
|          | URL | <kbd>Shift</kbd> + <kbd>Alt</kbd>+<kbd>U</kbd> |



- [`statcheck`](http://statcheck.io): Extract Statistics from Articles and Recompute p Values
- [`retractcheck`](https://frederikaust.shinyapps.io/retractcheck_shinyapp/): Check DOIs in a paper for retractions


## Reproducible software environment

- sessionInfo()
- packrat
- checkpoint
- liftr & docker

## RStudio

1. Document outline

RStudio provides a handy document outline view


<!-- If you discover an error in a data file, never modify it directly, but instead write code to correct the value. Explain why you made the fix. -->

<!-- Before you finish for the day, make sure you can knit the notebook (if you’re using caching, make sure to clear the caches). That will let you fix any problems while the code is still fresh in your mind. -->


## Splitting an R Markdown document

Some authors may prefer to split long manuscripts into multiple component files for better clarity.
There are two basic strategies to split R Markdown documents that can be combined or used in isolation: sourcing R scripts and splitting the R Markdown document.
If the R Markdown document contains a lot of code, it may be helpful to disincorporate parts of the code, such as reading, merging, restructuring, and relabeling data files.
The R scripts can then be executed at the respective section of the document using `source()`.

Some authors may prefer to split long manuscripts into a master or parent document and multiple children.
The master document, for example, consists of the YAML front matter and includes the children, which are themselves R Markdown documents without a YAML front matter.
To include a child document, insert an empty chunk and provide the path to the R Markdown document in the chunk option `child`.
    
It may be preferable to split long documents into multiple files

~~~
```{r child = "introduction.Rmd"}`r ''`
```

```{r child = "method.Rmd"}`r ''`
```

```{r child = "results.Rmd"}`r ''`
```

```{r child = "discussion.Rmd"}`r ''`
```
~~~

Search all files with <kbd>Ctrl</kbd> + <kbd>Shift</kbd>+<kbd>F</kbd>



## Workflow

- Don't edit raw data
- Don't set a working directory
- Generate graphics from R code
- Don't edit word documents
- Relative paths


## Best practices

1. Load all R packages in the first code chunk
    - Never include `install.packages()`
2. Set a seed for random number generators<br />(e.g., `set.seed()`)
3. Never use `setwd()`!
4. Use relative paths or load files from a permanent location
5. Use meaningful chunk names
6. Keep R code close to the corresponding prose
7. Document R and R-package versions<br />(e.g., `devtools::session_info()`)
8. Make sure you can knit without errors before going home



## Troubleshooting

### R

<!-- Copy and paste from R4DS -->
<!-- Troubleshooting R Markdown documents can be challenging because you are no longer in an interactive R environment, and you will need to learn some new tricks. The first thing you should always try is to recreate the problem in an interactive session. Restart R, then “Run all chunks” (either from Code menu, under Run region), or with the keyboard shortcut Ctrl + Alt + R. If you’re lucky, that will recreate the problem, and you can figure out what’s going on interactively. -->

<!-- If that doesn’t help, there must be something different between your interactive environment and the R markdown environment. You’re going to need to systematically explore the options. The most common difference is the working directory: the working directory of an R Markdown is the directory in which it lives. Check the working directory is what you expect by including getwd() in a chunk. -->

<!-- Next, brainstorm all the things that might cause the bug. You’ll need to systematically check that they’re the same in your R session and your R markdown session. The easiest way to do that is to set error = TRUE on the chunk causing the problem, then use print() and str() to check that settings are as you expect. -->


Fixing bugs in R Markdown documents can be challenging

- The code is run in a new non-interactive R session
- Try is to recreate the problem in your interactive session
    1. Restart R (`Session > Restart R` or<br /><kbd>Ctrl</kbd> + <kbd>Shift</kbd> + <kbd>F10</kbd>)
    2. Run all chunks (<kbd>Ctrl</kbd> + <kbd>Alt</kbd> + <kbd>R</kbd>)
    3. Compare the working directories (e.g., use `getwd()` in a code chunk)
    

