--- 
title           : "papaja: Reproducible APA manuscripts with R Markdown"
author          : "Frederik Aust & Marius Barth"
institute       : "University of Cologne, Germany"
date            : "`r Sys.Date()`"

description     : >
  ""

github-repo     : crsh/papaja_man

bibliography    : [references.bib, r-references.bib]
csl             : apa6.csl
link-citations  : yes

documentclass   : book
site            : bookdown::bookdown_site
output          :
  bookdown::gitbook:
    includes:
      after_body: footer.html

---

```{r libraries_and_options, echo = FALSE, message = FALSE, warning = FALSE}
library("rmarkdown")
library("car")
# library("ez")

library("papaja")
library("DiagrammeR")
library("ggplot2")

knitr::opts_chunk$set(fig.align = "center")
```

```{r cite_r, echo = FALSE}
r_refs("r-references.bib")
r_citations <- cite_r("r-references.bib", footnote = FALSE, pkgs = c("Matrix", "estimability", "reshape2"))
```

```{r not_on_cran, echo = FALSE}
not_on_cran <- "`papaja` has not yet been submitted to CRAN because it is under active development.
Currently, there are still a couple of loose ends we would like to tie up before we release the package to a larger audience.
If you would like to contribute to speed up the process, have a look at the chapters [Limitations] and [Future directions]."
```



# Preface {-}

This book will be a manual for the R package [papaja](https://github.com/crsh/papaja).
It is currently in the process of being written; I'm publishing new sections of the book as I'm writing them.
If you have suggestions for improvements or additional topics you would like to see covered, please let me know by opening an [issue on GitHub](https://github.com/crsh/papaja_man/issues) or creating a pull request.

`r not_on_cran`

If you are interested in reproducible research with R more generally, I recommend the books by @R-knitr and @gandrud_reproducible_2013.
Furthermore, @stodden_implementing_2014 give an overview of tools, practices, and platforms to implement reproducible research.

This manual is fully reproducible and was written in `r r_citations`.
Icons are provide royalty

---

```{r session_info, echo = FALSE}
devtools::session_info()
```

